\section{Relations}

We will first state the significant properties relations
can have. Taking a relation $R$ on $X$ with $x$, $y$, $z$
arbitrary in $X$: \begin{center}
    \begin{tabular}{ c l }
        \textbf{Name} & \textbf{Property} \\
        \hline
        Reflexive & $xRx$ \\
        Irreflexive & $\neg(xRx)$ \\
        \\
        Symmetric & $xRy \Rightarrow yRx$ \\
        Antisymmetric & $[xRy$ and $yRx] \Rightarrow [x = y]$ \\
        \\
        Connected & $[x = y]$ or $[xRy]$ or $[yRx]$ \\
        \\
        Transitive & $[xRy$ and $yRz] \Rightarrow [xRz]$ \\
    \end{tabular}
\end{center} For example, equivalence relations must satisfy 
reflexivity, symmetry, and transitivity.

\subsection{Partial Orderings}

We say that a relation $\prec$ on a set $X$ is a (strict) partial
ordering if it is irreflexive and transitive.
\\[\baselineskip]
Similarly, we say that a relation $\preceq$ on a set $X$ is a non-strict
partial ordering if it is reflexive, antisymmetric, and transitive.
\\[\baselineskip]
A partial ordering $(X, \prec)$ is wellfounded if for any non-empty subset
$Y$ of $X$, $Y$ has a least element under $\prec$.

\vfill

\subsection{Bounding}

For a partially ordered set $(X, \prec)$: \begin{itemize}
    \item $x_0$ in $X$ is the minimum of $X$ if for all $x$ in $X$,
        $x_0 \preceq x$,
    \item $x'$ in $X$ is minimal in $X$ if for all $x$ in $X$,
        $\neg(x \prec x')$,
    \item $x_1$ in $X$ is the maximum of $X$ if for all $x$ in $X$,
        $x \preceq x_1$,
    \item $x'$ in $X$ is maximal in $X$ if for all $x$ in $X$,
        $\neg(x' \prec x)$.
\end{itemize} Taking a non-empty subset $Y$ of $X$, we consider
the subordering $(Y, \prec)$ and for some $\alpha$ in $X$ we say:
\begin{itemize}
    \item $\alpha$ is a lower bound for $Y$ if for all $y$ in $Y$,
        $\alpha \prec y$,
    \item $\alpha$ is the infimum of $Y$ if it's a lower bound and
        for all lower bounds $\lambda$ of $Y$, $\alpha \preceq \lambda$,
    \item $\alpha$ is an upper bound for $Y$ if for all $y$ in $Y$,
        $y \prec \alpha$,
    \item $\alpha$ is the supremum of $Y$ if it's an upper bound and
        for all upper bounds $\tau$ of $Y$, $\tau \preceq \alpha$.
\end{itemize}

\subsection{Order Preserving Maps}

We say that $f : (X, \prec_1) \to (Y, \prec_2)$ is an order
preserving map if for each $x_1$, $x_2$ in $X$: \begin{align*}
    x_1 \prec_1 x_2 \Longrightarrow f(x_1) \prec_2 f(x_2).
\end{align*} Two orderings are (order) isomorphic if there is
a bijective order preserving map between them.

\subsection{Representation Theorem for Partially Ordered Sets}

For a partially ordered set $(X, \prec)$, there is a set 
$Y \subseteq \mathcal{P}(X)$ which is such that $(X, \preceq)$
is order isomorphic to $(Y, \subseteq)$.

\begin{proof}
   For some $x$ in $X$, we set $X^x = \{x' \in X : x' \preceq x\}$,
   the set of elements preceding or equal to $x$. For $x$, $y$
   in $X$, $x \neq y$ implies that $X^x \neq X^y$ as these sets
   contain $x$ and $y$ (resp.) so $x \mapsto X^x$ is injective.
   This map is surjective trivially (mapping from $X$ to
   $\{X^x : x \in X\}$).
   We have that: \begin{align*}
       x \preceq y \Longleftrightarrow X^x \subseteq X^y,
   \end{align*} by our definition. Thus, $x \mapsto X^x$ is
   an order isomorphism.
\end{proof}

\subsection{Total Orderings}

A relation $\prec$ on a set $X$ is a (strict) total ordering if it
is a connected strict partial ordering.
\\[\baselineskip]
Similarly, we say that a relation $\preceq$ on a set $X$ is a non-strict
total ordering if it is a connected non-strict partial ordering.

\subsection{Well-orderings}

A relation $\prec$ on a set $X$ is a well-ordering if it is a
strict total ordering and for any non-empty subset $Y$ of $X$,
$Y$ has a least element under $\prec$. We denote this with 
$(X, \prec) \in WO$.

\subsection{Ordered Pairs}

For $x$, $y$ sets, the ordered pair of $x$ and $y$ is the set: \begin{align*}
    \langle x, y \rangle = \{\{x\}, \{x, y\}\}.
\end{align*}

\subsubsection{Uniqueness of Ordered Pairs}

For $x$, $y$, $u$, $v$ sets, we have that: \begin{align*}
    \langle x, y \rangle = \langle u, v \rangle 
    \Longleftrightarrow
    (x = u) \text{ and } (y = v).
\end{align*}

\begin{proof}
    Suppose the former, if $x = y$ then 
    $\langle x, y \rangle = \{\{x\}, \{x, x\}\} = \{\{x\}\}$.
    Thus, $\langle u, v \rangle = \{\{u\}\}$ as it is equal to
    $\langle x, y \rangle$ which has one element, hence $u = v$. 
    By the Axiom of Extensionality, we have that $x = u$ and so
    $y = x = u = v$.
    \\[\baselineskip]
    If $x \neq y$, then $\langle x, y \rangle$ and $\langle u, v \rangle$
    both have the same two elements by our assumption (so $u \neq v$).
    We cannot have $\{x\} = \{u, v\}$ so $\{x\} = \{u\}$ which means
    $x = u$ by the Axiom of Extensionality. Thus, 
    $\{u, v\} = \{x, y\} = \{u, y\}$ so $y = v$.
    \\[\baselineskip]
    Suppose the latter, then the former holds trivially.
\end{proof}

\subsubsection{The Ordered $k$-tuple}

We define the $k$-tuple inductively. The $2$-tuple is already defined.
We define the $3$-tuple: \begin{align*}
    \langle x_1, x_2, x_3 \rangle &= \langle \langle x_1, x_2 \rangle, x_3 \rangle,
\end{align*} and for $k$ in $\{3, 4, \ldots\}$: \begin{align*}
    \langle x_1, x_2, \ldots, x_k \rangle &= \langle \langle x_1, x_2, \ldots, x_{k - 1} \rangle, x_k \rangle.
\end{align*}

\subsubsection{The Product of Sets}

For $A$, $B$ sets, we define: \begin{align*}
    A \times B = \{\langle a, b \rangle : a \in A, b \in B\}.
\end{align*} Similarly to $k$-tuples, for $A_1, A_2, \ldots, A_k$ sets,
we have $A_1 \times A_2$ defined, so we define: \begin{align*}
    A_1 \times A_2 \times \cdots \times A_k 
    &= (A_1 \times A_2 \times \cdots \times A_{k - 1}) \times A_k,
\end{align*} defining the $k$-product for $k$ in $\{2, 3, \ldots\}$.
This is not associative.

\subsection{Binary Relations}

A binary relation $R$ is a class of ordered pairs.
We write $R^{-1} = \{\langle y, x \rangle : \langle x, y \rangle \in R\}$.

\subsubsection{Domain and Range}

For a relation $R$, we define: \begin{align*}
    \Dom(R) &= \{x : \exists \, y \text{ where } \langle x, y \rangle \in R\}, \\
    \Ran(R) &= \{y : \exists \, x \text{ where } \langle x, y \rangle \in R\}, \\
    \Field(R) &= \Dom(R) \cup \Ran(R).
\end{align*}

\subsection{Functions}

A relation $F$ is a function if for all $x$ in $\Dom(F)$, there is a unique
$y$ in $\Ran(F)$ with $\langle x, y \rangle$ in $F$.
\\[\baselineskip]
If $F$ is a function, it is injective if and only if for all $x$, $x'$:
\begin{align*}
    (\langle x, y \rangle \in F \text{ and } \langle x', y \rangle \in F) 
    \Rightarrow
    (x = x').
\end{align*} 

\subsubsection{Ranges and Restrictions}

For $F : X \to Y$: \begin{itemize}
    \item $F''A = \{y \in Y : \exists \, x \in A \text{ such that } F(x) = y\}$
        the range of $F$ on $A$,
    \item $F \upharpoonright A = \{\ang{x, y} \in F : x \in A\}$
        the restriction of $F$ to $A$.
\end{itemize} We can see that $F''A = \Ran(F \upharpoonright A)$.

\subsubsection{The Set of Functions}

For $X$, $Y$ sets, we have that ${}^XY = \{F : F : X \to Y\}$.

\subsubsection{Indexed Cartesian Products}

For a set $I$ with each $i$ in $I$ corresponding to a
non-empty set $A_i$: \begin{align*}
    \prod_{i \in I} A_i = 
    \{\text{functions } f : \Dom(f) = I \text{ and } f(i) \in A_i \text{ for all } i \in I\}.
\end{align*}
