\section{Complexity Theory}

\subsection{Markov's Inequality}

For $X \geq 0$ a random variable with mean $\mu$, for all
$t$ in $\mathbb{R}_{\geq 0}$: \begin{gather*}
  \mathbb{P}(X \geq t) \leq \frac{\mu}{t}.
\end{gather*} Thus, if we expect our algorithm to run in time $T$,
then for some $k$ in $\mathbb{R}_{> 0}$, the probability it
takes more than $kT$ is $\frac{1}{k}$.

\subsubsection{Expected Runtime}

If we have an algorithm $A$ and a fixed input $x$ with expected runtime
$T$, we run $A(x)$ and if it fails to terminate with $2T$. The probability
it fails $k$ in $\mathbb{Z}_{>0}$ many times in a row is $2^{-k}$.

\subsection{Decision Problems}

A decision problem is a problem such that the answer is in the set $\{\texttt{Yes},
\texttt{ No}\}$.

\subsection{Oracles}

An oracle is a construct that given its corresponding problem, solves it in
$O(1)$ time.

\subsection{Cook Reductions}

For the decision problems $X$, $Y$ with $O_Y$ the oracle for $Y$,
we have that a Cook reduction from $X$ to $Y$ is an algorithm $A_X$
which given an input of size $n$ runs in $O(x^n)$ time and makes
$O(x^n)$ calls to $O_Y$ (all with inputs of size $O(x^n)$).

\newpage

\subsubsection{Properties of Cook Reductions}

We have that for the decision problems $X, Y, Z$: \begin{itemize}
  \item $X \leq_c Y$, $Y \leq_c Z$ implies that $X \leq_c Z$ (transitivity)
  \item $X \leq_c Y$ implies that if we have a polynomial time algorithm for $Y$, 
  we have one for $X$
  \item $X \leq_c Y$ and there is no polynomial-time algorithm for $X$ implies 
  that there is no polynomial-time algorithm for $Y$.
\end{itemize}

\subsection{Karp Reductions}

For the decision problems $X$ and $Y$, a Karp reduction from $X$ to $Y$
is a map $f$ from the instances of $X$ to the instances of $Y$ such that:
\begin{itemize}
  \item $f(x)$ can be computed in polynomial time (in $x$)
  \item $f(x)$ is a \texttt{Yes} instance of $Y$ if and only if it's a \texttt{Yes} instance
  of $X$.
\end{itemize} This is denoted by $X \leq_K Y$.

\subsubsection{Properties of Karp Reductions}

We have that Karp reductions are stronger than Cook reductions so that for any two
decision problems $X, Y$: \begin{gather*}
    X \leq_K Y \Rightarrow X \leq_C Y.
\end{gather*}

\subsection{The Class, \textbf{P}}

We have that \textbf{P} is the class of all decision problems which have a polynomial-time
algorithm. We have that $\textbf{P} \subseteq \textbf{NP}$ as we can just process the solution
(ignoring the witness).

\subsection{The Class, \textbf{NP}}

We have that \textbf{NP} is the class of all decision problems $X$ such that there is some
polynomial-time verification algorithm $A_X$ such that for some input $x$, 
if $x$ is a \texttt{Yes} instance of our problem, there is a witness bit string $w$
such that $A_X(x, w) = \texttt{Yes}$.

\subsubsection{\textbf{NP}-hardness}

We say a problem is \textbf{NP}-hard (under Cook reductions) if SAT Cook-reduces to it.
Similarly, we say a problem is \textbf{NP}-hard (under Karp reductions) if SAT Karp-reduces to it.

\subsubsection{\textbf{NP}-completeness}

We say a problem is \textbf{NP}-complete if it is \textbf{NP}-hard and in \textbf{NP}.

\subsection{The SAT Problem}

The SAT problem is the problem that asks if when given some formula in conjunctive-normal form
(consisting of AND and OR clauses) we can assign the variables such that the formula is 
satisfied (true).

\subsubsection{Cook-Levin Theorem}

We have that \textbf{SAT} is \textbf{NP}-hard and thus, \textbf{NP}-complete.
Thus, every problem which \textbf{SAT} reduces to is \textbf{NP}-hard.

\subsection{The 3-SAT Problem}

We have that the width of a conjunctive-normal form is the number of literals
of all the OR clauses. The 3-SAT problems asks if a width-3 conjunctive-normal 
form is satisfiable. We have that this is \textbf{NP}-complete.

\subsection{Further \textbf{NP}-complete Problems}

The following are \textbf{NP}-complete: \begin{itemize}
  \item Finding an independent set in a graph,
  \item Finding a vertex cover in a graph.
\end{itemize}