\section{Bilinear and Quadratic Forms}

\subsection{Bilinear Forms}

For $V$ a vector space over a field $K$, a bilinear form on $V$ is a map $\langle\,,\rangle : V \times V \to K$
such that: \begin{align*}
  \langle au + bv, w \rangle &= a \cdot \langle u, w \rangle + b \cdot \langle v, w \rangle \\
  \langle u, av + bw \rangle &= a \cdot \langle u, v \rangle + b \cdot \langle u, w \rangle,
\end{align*} for all $a, b$ in $K$, $u, v, w$ in $V$. Additionally, $\langle\,,\rangle$ is symmetric if 
$\langle u, v \rangle = \langle v, u \rangle$.

\subsection{Quadratic Forms}

For $V$ a vector space over a field $K$ with $\langle \, , \rangle$ a symmetric bilinear form on
$V$. The quadratic form $Q : V \to K$ associated to $\langle \, , \rangle$ is 
$Q(v) := \langle v, v \rangle$.

\subsubsection{Determining Bilinear Forms from Quadratic Forms}

We have that if char$(K) \neq 2$, then $\langle \, , \rangle$ is uniquely defined by $Q$ as:
\begin{align*}
  2^{-1}[Q(v + w) - Q(v) - Q(w)] 
  &= 2^{-1}[\langle v + w, v + w \rangle - Q(v) - Q(w)] \\
  &= 2^{-1}[\langle v, v + w \rangle + \langle w, v + w \rangle - Q(v) - Q(w)] \\
  &= 2^{-1}[2 \cdot \langle v, w \rangle] 
  \tag{$\langle \, , \rangle$ is symmetric} \\
  &= \langle v, w \rangle.
\end{align*} 

\subsection{Orthogonality}

Let $\langle \, , \rangle$ be a bilinear form on $V$ with $v, u$ in $V$. We say that $v$ in $V$ is 
orthogonal to $u$ if $\langle u, v \rangle = 0$. Note, be very careful as the bilinear is not
necessarily symmetric.

\subsubsection{Orthogonal Spaces}

For $W \subseteq V$, we have that $W^\perp$ is defined as: \begin{gather*}
  W^\perp = \{v \in V : w \in W, \langle w, v \rangle = 0\},
\end{gather*} the set of vectors such that for all $v$ in $W^\perp$, $v$ is orthogonal
to all of $W$. This is a subspace of $V$.

\subsubsection{The Kernel for Bilinear maps}

The kernel of $\langle \, , \rangle$ is $V^\perp$. If the kernel is $\{0_V\}$,
then the form is called non-degenerate and is called degenerate otherwise.

\subsubsection{Dimension and Orthogonal Spaces}

We have that if $V$ is finite dimensional and $\langle \, , \rangle$ is non-degenerate
then: \begin{gather*}
  \text{dim}(W^\perp) + \text{dim}(W) = \text{dim}(V).
\end{gather*}
\begin{proof}
    We note that $f$, the map corresponding to $\langle \, , \rangle$
    is an isomorphism. Thus: \begin{gather*}
        \Dim(W^\perp) = \Dim(f(W^\perp)) = \Dim(W^0) = \Dim(V) - \Dim(W). \qedhere
    \end{gather*}
\end{proof}

\subsection{Linear Maps from Bilinear Forms}

We can form a linear map $f : V \to V^*$ from a bilinear form $\langle \, , \rangle$ 
as follows: \begin{gather*}
  f(v)(u) = \langle u, v \rangle.
\end{gather*}

\subsubsection{Isomorphismic Bilinear Maps}

If $V$ is finite dimensional, we have that $f$ is an isomorphism if and only if 
$\langle \, , \rangle$ is non-degenerate. That is, $\text{Ker}(f) 
= \text{Ker}(\langle \, , \rangle) = V^\perp = \{0_V\}$.

\subsection{Matrices from Bilinear Forms}

For $V$ a finite $n$-dimensional vector space over $K$ with $S = \{v_1, \ldots, v_n\}$
an ordered basis for $V$. Let $B = \langle \, , \rangle$ be a bilinear form. The matrix corresponding
to $B$ with respect to $S$ is $M_{SS}(B) = (b_{ij})$ where: \begin{gather*}
  b_{ij} = \langle v_i, v_j \rangle.
\end{gather*} Similarly, taking $S^* = \{v^*_1, \ldots, v^*_n\}$ to be a dual basis for
$S^*$, we have a matrix $M_{SS^*}(f) = M_{SS}(B)$ for the linear map $f$ corresponding to $B$.

\subsubsection{Determining Bilinear Forms from Matrices}

Take $u, v$ in $V$ decomposed into vectors in $S$ with coefficients $x_1, \ldots, x_n$
and $y_1, \ldots, y_n$ respectively. Thus: \begin{gather*}
  \langle u, v \rangle = (x_1, \ldots, x_n) \cdot M_{SS}(B) \cdot \begin{pmatrix}
    y_1 \\ \vdots \\ y_n
  \end{pmatrix}.
\end{gather*}

\subsubsection{Properties of Bilinear Matrices}

We have that: \begin{itemize}
  \item $M_{SS}(B)$ is symmetric if and only if $B$ is symmetric,
  \item $B$ is non-degenerate if and only if $M_{SS}(B)$ is invertible.
\end{itemize}

\subsubsection{Similarity of Matrices of Bilinear Forms}

For $V$ a finite $n$-dimensional vector space over the field $K$ with char$(K) \neq 2$, 
let $S = \{v_1, \ldots, v_n\}$, $S' = \{v_1', \ldots, v_n'\}$ be
ordered bases for $V$. Let $B = \langle \, , \rangle$ be a symmetric bilinear form.
Let $C = C_{SS'}$ be the transition matrix. We have that: \begin{gather*}
  M_{S'S'}(B) = C^tM_{SS}(B)C.
\end{gather*}

\subsubsection{Diagonal Matrices of Bilinear Forms}

For $V$ a finite $n$-dimensional vector space over the field $K$ with char$(K) \neq 2$, 
let $B = \langle \, , \rangle$ be a symmetric bilinear form.
There exists a basis $S = \{v_1, \ldots, v_n\}$ for $V$ consisting of pairwise orthogonal
vectors and thus, the matrix $M_{SS}(B)$ is diagonal.

\subsection{Inner Products}

For a vector space $V$ over $K$ with symmetric bilinear form $B : V \times V \to K$, we
have that $B$ is an inner product if for all $v$ in $V$: \begin{gather*}
  B(v, v) \geq 0,
\end{gather*} and $B(v, v) = 0$ if and only if $v = 0_V$.